Graph Topological Aspects of Granger Causal Network Learning

### 2019-07-09 paper/paper.tex
Hacking away at classifying graphs...  A / C is extremely difficult, but there appear to be much clearer separations between individuals.

\section{Application}
\label{sec:application}
\section{Conclusion}
\label{sec:conclusion}
In this paper we have argued that considering particular topological
properties of Granger-causality networks can provide substantial
theoretical insights as well as computational benefits.  In
particular, the notion of a strongly-causal graph has been exploited
to establish conditions under which pairwise causality testing alone
is sufficient for recovering a complete Granger-causality graph.
Moreover, examples from the literature suggest that such topological
assumptions may be reasonable in some applications and secondly, even
when the strong-causality assumption is not met, we have provided some
empirical evidence to suggest that our pairwise testing algorithm PWGC
can still outperform the LASSO and adaLASSO, both of which are
commonly employed in applications.  Application to more general graphs
may also be within reach via an iterative graph building process,
i.e. estimating a sequence of SCGs via PWGC applied to the residuals
of the previous model.

>>>>

\section{Application}
\label{sec:application}
In this section we apply our methods to a real set of EEG data
obtained from the ``EEG Database Data
Set''\footnote{http://archive.ics.uci.edu/ml/datasets/EEG+Database}
\cite{zhang1995event} on the UCI machine learning repository
\cite{uci_mlr}.  This dataset contains 1 second long measurements of
(64 channel) EEG signals from patients who are given visual stimuli
and are broken into two groups: control and alcoholic.  The dataset
then consists of
$\mathcal{D} = \{\big(x^{(i)}(t)\big)_{t = 1}^T, y^{(i)} \}_{i = 1}^N$
where $T = 256$, $x^{(i)}(t) \in \R^{64}$,
$y^{(i)} \in \{\text{A}, \text{C}\}$, and \todo{fill in N} $N = ????$.

We have constructed a simple pipeline for classifying each each trial
as either ``A'' or ``C'' by first applying an iterative PWGC algorithm
(see Algorithm \ref{alg:iterated_pwgc}) directly to the EEG data
$x^{(i)}$ to obtain a Granger-causality graph $\widehat{G}^{(i)}$ then
simply feeding the vectorized adjacency matrix of $\widehat{G}^{(i)}$
through a Kernel PCA stage and then a Support Vector Machine with
parameters fit by cross validation.

While it is almost certainly possible to achieve much greater
classification accuracy on this dataset by constructing a classifier
to act directly on $x^{(i)}(t)$, our purpose is to demonstrate that
relevant information is being captured by the graph estimates
$\widehat{G}^{(i)}$.

Figures \ref{fig:ar_eeg_example} and \ref{fig:eeg_oos_eeror}
illustrate the appropriateness of modelling $x(t)$ with $\VAR$ models,
as opposed simply to a collection of $n$ unidimensional
autoregressiive models.  The conclusion being that the
Granger-causality graph estimates are not purely spurious.

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \caption{$\mathsf{AR}(p)$ Model Example}
    \label{fig:ar_eeg_example}
    \includegraphics[width=\linewidth]{eeg_estimate_example.pdf}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \caption{OOS Error for a Single Subject}
    \label{fig:eeg_oos_error}
    \includegraphics[width=\linewidth]{eeg_oos_demonstration.pdf}
  \end{subfigure}

  {\scriptsize Figure \ref{fig:ar_eeg_example} provides an example of an
  autoregressive model fir via the Levinson-Durbin algorithm (choosing
  the system order via the BIC) on a single EEG channel.  Figure
  \ref{fig:eeg_oos_error} demonstrates the improvement of a single
  $\VAR$ model over independent $\mathsf{AR}$ models.

  Out-of-Sample estimates are performed on data that were not used in
  fitting the models.}

\end{figure}

\section{Conclusion}
\label{sec:conclusion}
In this paper we have argued that considering particular topological
properties of Granger-causality networks can provide substantial
theoretical insights as well as computational benefits.  In
particular, the notion of a strongly-causal graph has been exploited
to establish conditions under which pairwise causality testing alone
is sufficient for recovering a complete Granger-causality graph.
Moreover, examples from the literature suggest that such topological
assumptions may be reasonable in some applications and secondly, even
when the strong-causality assumption is not met, we have provided some
empirical evidence to suggest that our pairwise testing algorithm PWGC
can still outperform the LASSO and adaLASSO, both of which are
commonly employed in applications.  Application to more general graphs
may also be within reach via an iterative graph building process,
i.e. estimating a sequence of SCGs via PWGC applied to the residuals
of the previous model.
